Covariates analysis of MSBB RNASeq data of region BM 44

### 2016-05-01 code/Rmd/MSSM_BM_44_RNASeq.Rmd
Minor bug fix

rownames(COVARIATES) <- rownames(METADATA)

# Sub sample the analysis
COVARIATES = rownameToFirstColumn(COVARIATES, 'ID')
COVARIATES = COVARIATES %>% 
  group_by(BrainRegion.CDR, SEX, RACE, Batch) %>% 
  top_n(2) %>%
  data.frame
COUNT = COUNT[,COVARIATES$ID]
rownames(COVARIATES) = COVARIATES$ID
COVARIATES$ID = NULL
```
Covariate correlation
```{r covariates.correlation, fig.width=10, fig.height=10}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
ggheatmap.show(COVARIATES.CORRELATION$plot, col.width=0.3, row.width=0.15)
```
### CPM Normalisation
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=BrainRegion.CDR, shape=SEX, size=PMI))
p <- p + theme_bw() + theme(legend.position="top") + scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= ID), size=4, hjust=1)
p

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Batch), shape=as.factor(CDR), size=as.numeric(PMI)))
p <- p + theme_bw() + theme(legend.position="top") + scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
p

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = rownames(RESIDUAL.GENE_EXPRESSION),
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id"),
                                      Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$CDR))
colnames(tmp) = 'No. of samples'
kable(tmp)
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=18}
# Fit contrast
contrast = colnames(VOOM.ADJUSTED.FIT$coefficients) %>%
  grep('Dx',.,value=T) %>%
  combn(2) %>%
  apply(2,paste,collapse='-') %>%
  makeContrasts(contrasts=.,
                levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT = eBayes(FIT)

>>>>

rownames(COVARIATES) <- rownames(METADATA)
```
Covariate correlation
```{r covariates.correlation, fig.width=10, fig.height=10}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
ggheatmap.show(COVARIATES.CORRELATION$plot, col.width=0.3, row.width=0.15)
```
### CPM Normalisation
Preprocess counts matrix and metadata. Determine design matrix for normalisation and differential expression analysis.

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=Dx, shape=SEX, size=PMI))
p <- p + theme_bw() + theme(legend.position="top") + scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= ID), size=4, hjust=1)
p

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=factor(Batch), shape=as.factor(Dx), size=as.numeric(PMI)))
p <- p + theme_bw() + theme(legend.position="top") + scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= ID), size=4, hjust=0)
p

@@

# Query biomart
Ensemble2HGNC <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"),
                       filters = "ensembl_gene_id", values = rownames(RESIDUAL.GENE_EXPRESSION),
                       mart = mart)
RESIDUAL.GENE_EXPRESSION <- left_join(rownameToFirstColumn(RESIDUAL.GENE_EXPRESSION, "ensembl_gene_id"),
                                      Ensemble2HGNC) %>%
  dplyr::select(ensembl_gene_id, hgnc_symbol, one_of(rownames(COVARIATES)))
```
```{r summary, results='asis', echo=FALSE}
writeLines('Number of samples in each category are')
tmp = as.data.frame(summary(COVARIATES$Dx))
colnames(tmp) = 'No. of samples'
kable(tmp)
```
### Differential expression analysis
Genes that are differentially expressed at an FDR <= 0.05 are
```{r diffExp, fig.height=10, fig.width=18}
# Fit contrast
contrast = colnames(VOOM.ADJUSTED.FIT$coefficients) %>%
  grep('Dx',.,value=T) %>%
  combn(2) %>%
  apply(2,paste,collapse='-') %>%
  makeContrasts(contrasts=.,
                levels = colnames(VOOM.ADJUSTED.FIT$coefficients))
FIT = contrasts.fit(VOOM.ADJUSTED.FIT, contrasts=contrast)
FIT = eBayes(FIT)
